Capstone Movielens Report

### 2025-12-09 reports/capstone-movielens.site/05.appendix.Rmd
Appendix A review, function refined: mean_reg.

```{r eval=FALSE}
mean_reg(vals, lambda = 0, na.rm = TRUE)

@@

:::
  - **na.rm:** a logical evaluating to TRUE or FALSE indicating whether NA values should be stripped before the computation proceeds.

@@

The function calculates the _penalized mean_ of the values passed in the [vals](#func.mean_reg.args.vals) argument as follows:
$$
\mu(\lambda) = \frac{1}{\lambda + N} \sum_{i=1}^{N} x_i
$$

@@

We call this function internally from the functions that compute the _penalized estimates_  as described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the *Course Textbook (New Edition)*[@IDS2_24-2].

@@

The [penalized mean](#func.mean_reg.details) of the values passed in the [vals](#func.mean_reg.args.vals) argument.
:::

@@

The soure code of the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L70) function is shown below:
```{r eval=FALSE}
mean_reg <- function(vals, lambda = 0, na.rm = TRUE){
  if (is.na(lambda)) {
    stop("Function: mean_reg
`lambda` is `NA`")
  }

@@

```{r eval=FALSE}
  train_set |>
    calc_date_general_effect(lambda) |>
    calc_UMGY_SmoothedDay_effect()
```

@@

The source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect <- function(date_general_effect){
  date_general_effect |>
    group_by(year) |>
    summarise(ye = mean(de, na.rm = TRUE))
}
```

@@

```{r eval=FALSE}
    cv_fold_dat$validation_set |> UMGY_SmoothedDay_effect.predict(UMGYD_effect)
```

@@

The source code of the [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L125) function is shown below:
```{r eval=FALSE}
UMGY_SmoothedDay_effect.predict <- function(test_set, UMGYD_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(UMGYD_effect, by='year') |>
    mutate(resid = rating - clamp(mu + a + b + g + ye)) |> 
    pull(resid) |> mse()
}
```

@@

```{r eval=FALSE}
    cv_fold_dat$validation_set |> calc_UMGY_SmoothedDay_effect.MSE(UMGYD_effect)
```

@@

The source code of the [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L141) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE <- function(test_set, UMGYD_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(UMGYD_effect, by='year') |>
    mutate(resid = rating - clamp(mu + a + b + g + ye)) |> 
    pull(resid) |> mse()
}
```

@@

```{r eval=FALSE}
  UMGY_effect_RMSE <- sqrt(calc_UMGY_SmoothedDay_effect.MSE.cv(UMGYD_effect))
```

@@

The simplified version of the source code of the [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L147) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE.cv <- function(UMGYD_effect){
  start <- put_start_date()
  UMGY_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> calc_UMGY_SmoothedDay_effect.MSE(UMGYD_effect)
  })
  put_end_date(start)
  put_log1("Function: calc_UMGY_SmoothedDay_effect.MSE.cv
Date (Year) Effect MSE values have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

@@

```{r eval=FALSE}
cv.UMGYD_effect.RMSE <- calc_UMGY_SmoothedDay_effect.RMSE.cv(cv.UMGYD_effect)
```
:::

@@

The source code of the [calc_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L169) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.RMSE.cv <- function(UMGYD_effect){
  UMGY_effect_RMSE <- sqrt(calc_UMGY_effect_MSE.cv(UMGYD_effect))
  put_log2("%1-Fold Cross Validation ultimate RMSE: %2", 
           CVFolds_N, 
           UMGY_effect_RMSE)

@@

```{r eval=FALSE}
  lambdas <- seq(0, 512, 32)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, 
                     regularize.test_lambda.UMGYD_effect.cv,
                     steps.beyond_min = 16)
```

@@

The source code of the [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L130) function is shown below:
```{r eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGYD_effect.cv
`lambda` is `NA`")
  }

>>>>

```{r eval=FALSE}
  UM.effect <- train_set |>
    left_join(edx.user_effect, by = "userId") |>
    mutate(resid = rating - (mu + a)) |> 
    group_by(movieId) |>
    summarise(b = mean_reg(resid, lambda), n = n())
```

::: {.noteblock data-latex=""}
In the sample code snippet above, the `mean_reg` function is called internally from the [train_user_movie_effect](#func.train_user_movie_effect) function described below in section *Models Training: Support Functions \* [UME Model: Utility Functions] of *this Appendix*.
:::
:::

@@

:::
  - **na.rm:** (`TRUE` by default) Logical value indicating whether `NA` values should be stripped before the computation proceeds.

@@

This function is used in the *regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-2]) to calculate the _penalized mean_ of the values passed in the [vals](#func.mean_reg.args.vals) argument as follows:
$$
\mu(\lambda) = \frac{1}{\lambda + N} \sum_{i=1}^{N} x_i
$$

::: {.noteblock data-latex=""}
This is a helper function intended for use in other helper functions that compute *penalized estimates* during the *regularization* process of the models being trained[@IDS2_24-2]. All these functions are described below in section [Models Training: Support Functions] of *this Appendix*.
:::

@@

@@

(Numeric) *Penalized mean* of the values passed in the [vals](#func.mean_reg.args.vals) argument.
:::

@@

The source code of the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L70) function is shown below:
```{r eval=FALSE}
mean_reg <- function(vals, lambda = 0, na.rm = TRUE){
  if (is.na(lambda)) {
    stop("Function: mean_reg
`lambda` is `NA`")
  }

@@

@@

The source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L76) function is shown below:
```{r eval=FALSE}
```

@@

@@

The source code of the [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L125) function is shown below:
```{r eval=FALSE}
```

@@

@@

The source code of the [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L141) function is shown below:
```{r eval=FALSE}
```

@@

@@

The simplified version of the source code of the [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L147) function is shown below:
```{r eval=FALSE}
```

@@

@@

The source code of the [calc_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L169) function is shown below:
```{r eval=FALSE}
```

@@

@@

The source code of the [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L130) function is shown below:
```{r eval=FALSE}
```
